---
title: "FInalProject"
author: "Ryan Wissman"
date: "August 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Summary

This is the final assignment for the Coursera course titled "Practical Machine Learning." The goal of this project is to predict the manner in wich a participant did the exercise. We will accomplish this by using a random forest classifier. The Caret package will be utilized.

## Load the Data

First we will load the data from the corresponding CSV files. We will also load the appropriate packages.

```{r}
data_training <- read.csv("pml-training.csv")
data_testing <- read.csv("pml-training.csv")
```

#### Describe the data

```{r echo=TRUE}
dim(data_training)
names(data_training)[160]
```

After doing a cursory view of the data we can see that there are a total of 19622 observations with 160 variables, our target variable name is *classe*.

## Process the data

Before we start analyzing the data we need to remove NA (empty) values. 

```{r}

```
