---
title: "FInalProject"
author: "Ryan Wissman"
date: "August 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Summary

This is the final assignment for the Coursera course titled "Practical Machine Learning." The goal of this project is to predict the manner in wich a participant did the exercise. We will accomplish this by using a random forest classifier. The Caret package will be utilized.

## Load the Data

First we will load the data from the corresponding CSV files. We will also load the appropriate packages, in this case we will be using the caret package as well as the randomforest package. 

```{r}
#load packages
library(caret)
library(randomForest)

#load data
data_training <- read.csv("pml-training.csv", na.strings=c("", "NA", "NULL"))
data_testing <- read.csv("pml-testing.csv", na.strings=c("", "NA", "NULL"))
```

#### Describe the data

```{r echo=TRUE}
dim(data_training)
names(data_training)[160]
```

After doing a cursory view of the data we can see that there are a total of 19622 observations with 160 variables, our target variable name is *classe*.

## Process the data

Since there are also a lot of missing data in some of the variables we can remove them. We will do this for both the training set and test sets. 

```{r}
blanks <- apply(data_training, 2, function(counts) {sum(is.na(counts))})
data_training <- data_training[,which(blanks==0)]

#do the same for the original test set
data_testing <- data_testing[,which(blanks==0)]
```

The next thing we will do will be to remove irrelevant variables (columns) from our dataset. The variables that we will remove will be as follows:

* *x*, simply an index
* *user_name*, username
* *raw_timestamp_part_1* machine timestamp of record
* *raw_timestamp_part_2* machine timestamp of record
* *cvtd_timestamp*, when the record was taken in readable form

```{r}
data_training <- data_training[,-grep("^X|^user_name|^raw_timestamp_part_1
                                            |^raw_timestamp_part_2|^cvtd_timestamp", names(data_training))]

#do the same for the original test set
data_testing <- data_testing[,-grep("^X|^user_name|^raw_timestamp_part_1
                                            |^raw_timestamp_part_2|^cvtd_timestamp", names(data_testing))]
```

Now we will split the training data set into a 70/30 training and test set split (70% training, 30% test). 

```{r echo=TRUE}
inTrain <- createDataPartition(y=data_training$classe, p=0.7, list=FALSE)
training <- data_training[inTrain,]
testing <- data_training[-inTrain,]
dim(training)
dim(testing)
```

Now that we processed the data we can build a prediction model.

## Build the model

The prediction model will use the random forest method to predict the class of activity (exercise) performed. THis method can be a little slow, but produces a large random forest with 19 elements.

### Model Fit

Fit the prediction model.

```{r}
model_fit <- randomForest(classe ~ ., data = training)
```

###Cross-Validation

We will perform cross-validation on the test set. We should have a fairly small out of sample error as the random forest method should be very accurate. We will then output a confusion matrix of the relevant statistics.

```{r echo=TRUE}
predict_fit <- predict(model_fit, testing)
confusionMatrix(testing$classe, predict_fit)
```

Our model achieves 99.93% accuracy on the test set. 

###Apply the model to the original test set

Now we will apply the model to the original test set 

```{r}
data_testing_model <- predict(model_fit, data_testing)
data_testing_model
```
